\section{Calculus with Parametric Curves}
  We will mainly solve problems involving tangents, area, arc length, and surface area.

  \subsection*{Tangents}
    In the previous section, we saw that some curves defined by parametric equations $ x = f(t) $ and $ y = g(t) $ can also be expressed, by eliminating the parameter, in the form $ y = F(x) $. If we substitute $ x = f(t) $ and $ y = g(t) $ in the equation $ y = F(x) $, we get
      $$ g(t) = F(f(t)) $$
    If g, f, and F are differentiable, the Chain Rule gives
      $$ g'(t) = F'(f(t))f'(t) = F'(x)f'(t) $$
    If $ f'(t) \neq 0 $, we can solve for $F'(x)$:
    \begin{definition}
      The slope of the tangent to the parametric curve $y=F(x)$ is $F'(x)$.
      $$ F'(x) = \frac{g'(t)}{f'(t)} $$
    \end{definition}
    This enables us to find tangents to parametric curves without having to eliminate the parameter. We can rewrite the
    previous equation in an easily remembered form.
    \begin{definition}
      We can use this to find tangents to parametric curves without having to eliminate the parameter.
      $$ \frac{dy}{dx} = \frac{\frac{dy}{dt}}{\frac{dx}{dt}} \ \ \text{if} \ \ \frac{dx}{dt} \neq 0$$
      The curve has a
      \begin{itemize}
        \item horizontal tangent when $ \frac{dy}{dt} = 0 $ (provided that $\frac{dx}{dt} \neq 0$)
        \item vertical tangent when $ \frac{dx}{dt} = 0 $ (provided that $\frac{dy}{dt} \neq 0$)
      \end{itemize}
      This is useful when sketching parametric curves.
    \end{definition}
    \begin{definition}
      We can also find $ \frac{d^2 y}{dx^2}$ by replacing $y$ with $\frac{dy}{dx}$
      $$ \frac{d^2 y}{dx^2} = \frac{d}{dx}\left(\frac{dy}{dx}\right) = \frac{\frac{d}{dt}\left(\frac{dy}{dx}\right)}{\frac{dx}{dt}}$$
    \end{definition}
    \begin{minipage}{\linewidth}
      \begin{proof}
        Find $\frac{dy}{dx}$ and $\frac{d^2 y}{dx^2}$ considering $y(t)$ and $g(t)$.
        \begin{enumerate}
          \item \begin{equation*}
            \text{Chain rule: } \frac{dy}{dt} = \frac{dy}{dx}\frac{dx}{dt} \implies \frac{dy}{dx} = \frac{\frac{dy}{dt}}{\frac{dx}{dt}} \ \ \ (\implies \text{means "implies"})
          \end{equation*}
          \item \begin{align*}
            \text{Chain rule: } \frac{d}{dt}\left(\frac{dy}{dx}\right) &= \left(\frac{d}{dx}\frac{dy}{dx}\right)\frac{dx}{dt} = \frac{d^2 y}{dx^2}\frac{dx}{dt} \\
            \text{Substitute: } \frac{d}{dt}\left(\frac{dy}{dx}\right) &= \frac{d}{dt}\left(\frac{\frac{dy}{dt}}{\frac{dx}{dt}}\right) \\
            \text{Quotient rule: } &= \frac{\frac{d^2 y}{dt^2}\frac{d^x}{dt} - \frac{dy}{dt}\frac{d^2 x}{dt^2}}{\left(\frac{dx}{dt}\right)^2} \\
            \intertext{Set equation from line 1 and line 3 equal and divide both sides by $\frac{dx}{dt}$} \\
            \frac{d^2 y}{dx^2} &= \frac{\frac{d^2 y}{dt^2}\frac{d^x}{dt} - \frac{dy}{dt}\frac{d^2 x}{dt^2}}{\left(\frac{dx}{dt}\right)^2 \left(\frac{dx}{dt}\right)} \\
            &= \frac{\frac{d^2 y}{dt^2}\frac{d^x}{dt} - \frac{dy}{dt}\frac{d^2 x}{dt^2}}{\left(\frac{dx}{dt}\right)^3}
          \end{align*}
        \end{enumerate}
      \end{proof}
    \end{minipage}
    \begin{example}
      A curve C is defined by the parametric equations $ {x=t^2 ,\ y=t^3 -3t }$.
      \begin{enumerate}
        \item Show that $C$ has two tangents at the point (3,0) and find their equations.
        \item Find the points on $C$ where the tangent is horizontal or vertical.
        \item Determine where the curve is concave upward or downward.
      \end{enumerate}
    \end{example}
    \begin{solution}
      A curve C is defined by the parametric equations $ x=t^2 ,\ y=t^3 -3t $.
      \begin{enumerate}
        \item Rewrite $ y=t^3 -3t = t(t^2 -3) = 0 $ when $t=0$ or $t= \pm\sqrt{3} $. This indicates that C intersects itself at (3.0).
        \begin{align*}
          \frac{dy}{dx} &= \frac{dy/dt}{dx/dt} = \frac{3t^2 - 3}{2t} = \frac{3}{2}\left(t-\frac{1}{t}\right) \\
          t=\pm\sqrt{3} \rightarrow dy/dx &= \pm 6/(2\sqrt{3})
        \end{align*}
          so the equations of the tangents at (3,0) are
            $$ y = \sqrt{3}(x-3) \ \ \ and \ \ \ y = -\sqrt{3}(x-3) $$
        \item C has a horizontal tangent when $dy/dx=0$. In other words, when $dy/dt=0$ and $dx/dt \neq 0$. $dy/dt=3t^2 - 3 = 0$ when $t^2 = 1$ so $t=\pm 1$. This means there are horizontal tangents on C at (1,-2) and (1,2). C has a vertical tangent when $dx/dt = 2t = 0$, so $t= 0$. This means C has a vertical tangent at (0,0).
        \item To determine concavity we calculate the second derivative:
          $$ \frac{d^2 y}{dx^2} = \frac{\frac{d}{dt}\left(\frac{dy}{dx}\right)}{\frac{dx}{dt}} = \frac{\frac{3}{2}\left(1+\frac{1}{t^2}\right)}{2t} = \frac{3(t^2 + 1)}{4t^3} $$
          The curve is concave upward when $t>0$ and concave downward when $t<0$.
      \end{enumerate}
    \end{solution}

  \subsection*{Area}
    We already know that area under a curve $y=F(x)$ from $a$ to $b$ is $A=\int_{a}^{b} F(x)dx$. We can apply this to parametric equations using the Substitution Rule for Definite Integrals.
    \begin{definition}
      If the curve $C$ is given by parametric equations $x=f(t)$ and $y=g(t)$ and $t$ increases from $\alpha$ to $\beta$,
        $$ A=\int_{a}^{b} ydx = \int_{\alpha}^{\beta} g(t)f'(t)dt $$
      (Switch $\alpha$ to $\beta$ if the point on $C$ at $\beta$ is more left than $\alpha$.
    \end{definition}
    \begin{example}
      Find the area under one arch of the cycloid ${x=r(\theta-\sin\theta),\ y=r(1-\cos\theta)}.$
    \end{example}
    \begin{solution}
      One arch of the cycloid is given by $0 \leq \theta \leq 2\pi$. Using the Substitution Rule with $y=r(1-\cos\theta)$ and $dx=r(1-\cos\theta)d\theta$, we have
      \begin{align*}
        A &= \int_{0}^{2\pi} ydx = A=\int_{0}^{2\pi} r(1-\cos\theta)r(1-\cos\theta)d\theta \\
          &= r^2 \int_{0}^{2\pi} (1-\cos\theta)^2 d\theta = r^2 \int_{0}^{2\pi} (1 - 2 \cos\theta + \cos^2 \theta) d\theta \\
          &= r^2 \int_{0}^{2\pi} \left[1 - 2 \cos\theta + \frac{1}{2}(1+\cos 2\theta)\right] d\theta \\
          &= r^2 \left[ \frac{3}{2}\theta - 2 \sin\theta + \frac{1}{4}\sin2\theta \right]_{0}^{2\pi} \\
          &= r^2 \left( \frac{3}{2} \cdot 2\pi \right) = 3\pi r^2
      \end{align*}
    \end{solution}

  \subsection*{Arc Length}
    We already know how to find length $L$ of a curve $C$ given in the form ${y=F(x)}$, $a \leq x \leq b$.
    \begin{definition}
      If $F'$ is continuous, then $$L=\int_{a}^{b}\sqrt{1+\left(\frac{dy}{dx}\right)^2 dx}$$
    \end{definition}
    If $C$ can describe the parametric equations $x=f(t)$ and $y=g(t)$, $\alpha \leq t \leq \beta$, where $dx/dt = f'(t)>0$. Using the substitution rule, we obtain
      $$L=\int_{a}^{b}\sqrt{1+\left(\frac{dy}{dx}\right)^2 dx} = \int_{\alpha}^{\beta}\sqrt{1+\left(\frac{dy/dt}{dx/dt}\right)^2 \frac{dx}{dt} dt}$$
    Since $dx/dt > 0$, we have
    \begin{theorem}
      If a curve $C$ is described by the parametric equations $x=f(t)$, $y=g(t)$, $\alpha \leq t \leq \beta$, where $f'$ and $g'$ are continuous on $[\alpha,\beta]$ and $C$ is traversed exactly once as $t$ increases from $\alpha$ to $\beta$, then the length of $C$ is
        $$L = \int_{\alpha}^{\beta}\sqrt{\left(\frac{dx}{dt}\right)^2+\left(\frac{dy}{dt}\right)^2} dt$$
    \end{theorem}
    This is consistent with the general formula $L=\int ds$ and ${(ds^2)=(dx^2)+(dy^2)}$.
    \begin{proof}\let\qed\relax
      Prove the length formula of a parametric curve
      \begin{align*}
        \overrightarrow{ds} &= \overrightarrow{i}dx + \overrightarrow{j}dy\\
        ds^2 &= \overrightarrow{ds} \cdot \overrightarrow{ds} = \left(\overrightarrow{i}dx+\overrightarrow{j}dy\right) \cdot \left(\overrightarrow{i}dx+\overrightarrow{j}dy\right) = dx^2 + dy^2 \\
        ds &= \sqrt{dx^2 + dy^2} = \sqrt{\left(\frac{dx}{dt}\right)^2+\left(\frac{dy}{dt}\right)^2} dt \\
        L &= \int_{\alpha}^{\beta} ds = \int_{\alpha}^{\beta}\sqrt{\left(\frac{dx}{dt}\right)^2+\left(\frac{dy}{dt}\right)^2} dt
      \end{align*}
    \end{proof}
    \begin{example}
      Find the length of the unit circle as (x,y) moves both once and twice around the circle.
    \end{example}
    \begin{solution}
      For one traversal around the unit circle,
      $$ x=\cos t \ \ \ y=\sin t \ \ \ 0\leq t\leq 2\pi$$
      so $dx/dt = -\sin t$ and $dy/dt = \cos t$
      \begin{align*}
        L &= \int_{0}^{2\pi}\sqrt{\left(\frac{dx}{dt}\right)^2+\left(\frac{dy}{dt}\right)^2} dt = \int_{0}^{2\pi}\sqrt{\sin^2 t + \cos^2 t}\ dt \\
          &= \int_{0}^{2\pi}dt = 2\pi
      \end{align*}

      For two traversals around the unit circle,
      $$ x=\sin 2t \ \ \ y=\cos 2t \ \ \ 0\leq t\leq 2\pi$$
      so $dx/dt = 2\cos 2t$ and $dy/dt = -2\sin 2t$
      \begin{align*}
        L &= \int_{0}^{2\pi}\sqrt{\left(\frac{dx}{dt}\right)^2+\left(\frac{dy}{dt}\right)^2} dt = \int_{0}^{2\pi}\sqrt{4\cos^2 2t + 4\sin^2 2t}\ dt
          &= \int_{0}^{2\pi}2\ dt = 4\pi
      \end{align*}
    \end{solution}
  \subsection*{Surface Area}
    We can also adapt the surface area formula to a parametric curve.
    \begin{definition}
      If a curve $C$ is described by the parametric equations $x=f(t)$, $y=g(t)$, $\alpha \leq t \leq \beta$, is rotated about the \textbf{x-axis}, where $f'$,$g'$ are continuous and $g(t)\geq 0$, the surface area is
        $$ S=\int_{\alpha}^{\beta} 2\pi y \sqrt{\left(\frac{dx}{dt}\right)^2+\left(\frac{dy}{dt}\right)^2} dt  $$
      If the curve C is rotated about the \textbf{y-axis}, the surface area is
        $$ S=\int_{\alpha}^{\beta} 2\pi x \sqrt{\left(\frac{dx}{dt}\right)^2+\left(\frac{dy}{dt}\right)^2} dt  $$
      The generic formulas ${S=\int 2\pi y\ ds}$ for rotation about the x-axis and ${S=\int 2\pi x\ ds}$ for rotation about the y-axis are still valid, but for parametric curves we use
        $$ds = \sqrt{\left(\frac{dx}{dt}\right)^2+\left(\frac{dy}{dt}\right)^2} dt$$
    \end{definition}
    \begin{example}
      Show that the surface area of a sphere of radius $r$ is $4\pi r^2$
    \end{example}
    \begin{solution}
      The sphere is obtained by rotating the semicircle
        $$x=r\cos t \ \ \ y=r\sin t \ \ \ 0 \leq t \leq \pi$$
      about the x-axis.
      \begin{align*}
        S &= \int_{0}^{\pi} 2\pi r\sin t\sqrt{(-r\sin t)^2 + (r\cos t)^2}\ dt \\
          &= 2\pi\int_{0}^{\pi} r \sin t\sqrt{r^2 (sin^2 t + \cos^2 t)}\ dt \\
          &= 2\pi\int_{0}^{\pi} r \sin t \cdot r\ dt = 2\pi r^2\int_{0}^{\pi}\sin t\ dt \\
          &= 2\pi r^2 (-\cos t)]_{0}^{\pi} = 4\pi r^2
      \end{align*}
    \end{solution}

